---
title: "The fql package"
author: "jperreton@gmail.com"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fql}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, warning=FALSE, echo=FALSE, results='hide'}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning=FALSE, echo=FALSE, results='hide', message=FALSE}
library(fql)
```

This package is a collection of verbs that implement a DSL to work with dataframes. It is broadly inspired by [https://code.kx.com/v2/ref/]. FQL stands for Functional Query Language. This library has two dependencies:

* functional

* magrittr

### Meaty Dataframe queries

Let us create a dataset of daily price and volume for ORA and VOD. 
```{r meaty}
d <- data.frame(symbol=c(rep("ORA"  , 10), rep("VOD", 10)),
                date=as.character(rep(seq(as.Date("2019-01-01"),
                                          9+as.Date("2019-01-01"),by=1), 2)),
                price=c(14*cumprod(1+0.05*rnorm(10)), 150*cumprod(1+0.05*rnorm(10))),
                volume=c(cumsum(runif(10)), cumsum(runif(10)))
                  )
summary(d)

```

Now let us say that we want to compute the 10-day vwap and the average turnover by symbol:
```{r vwap}
d %>%
update(what=alist(turnover=price*volume)) %>%
select(by=alist(symbol=symbol),
       what=alist(turnover=mean(turnover),
                  vwap=sum(turnover) / sum(volume)))
```

Now let us say that we have the exchange in a separate table and we want to join it into the result:

```{r lj}
e <- data.frame(symbol=c("ORA", "VOD"), exchangeid=c(1, 34))
d %>%
update(what=alist(turnover=price*volume)) %>%
select(by=alist(symbol=symbol),
       what=alist(turnover=mean(turnover),
                  vwap=sum(turnover) / sum(volume))) %>%
lj(e %>% xkey("symbol"))
```

Now let us say we want to compute returns at various horizons. We can achieve it with metaprogramming:
```{r returnsexp}
n <- 1:5
r <- n %>% each(function(k) parse(text=paste("log(price/xprev(price, ", k, "))"))) %>%
     dict(paste0("r_", n), .)
head(r, 2)
```
Once we have prepared the returns expressions, we can then update d with it:
```{r returns}
d %>% 
update(by=alist(symbol=symbol), what=r) %>%
tail(5)
```

### Function Decorators

#### onCall

onCall allows decorating a function and execute a function call right at the start of its evaluation.
The default handler is to display its arguments.

```{r oncall}
f <- function(x) x
g <- onCall(f)
g(1)
```

#### onError

onError allows decorating a function and execute a handler whenever an error is thrown (tryCatch block).
Its default handler is to print the error and return NULL.

```{r onerror}
h <- onError(function(x) "1"-x)
h(1)
```

Notice now how the h function happily returns NULL rather than an error.
